---
title: "Day 6 Practice, Loops"
output: html_document
---

1. Create a chunk of code that does the following:
  - get a subset of data from the `iris` data set that just includes one species
  - get the mean of each of the values, and arrange these in a one-row data frame
  - make a PDF plot of Sepal.Width by Petal.Width

```{r}
myris <- iris
myris.virginica <- myris[myris$Species == "virginica", ]
sl <- mean(myris.virginica$Sepal.Length)
sw <- mean(myris.virginica$Sepal.Width)
pl <- mean(myris.virginica$Petal.Length)
pw <- mean(myris.virginica$Petal.Width)
my.answer <- data.frame(Species = "virginica", Sepal.Length = sl, Sepal.Width = sw, Petal.Length = pl, Petal.Width = pw)

# OR
myris <- iris
myris.virginica <- myris[myris$Species == "virginica", ]
myris.virginica.melt <- melt(myris.virginica, measure.vars = 1:4)
my.answer <- dcast(myris.virginica.melt, Species ~ variable, value = "value", fun.aggregate = mean)

```

2. Now make a copy of this code and turn it into a loop:
  - loop through all three species
  - combine the results from each loop into a single data frame
    - do this the "slow" way and the "fast" way
  - combine the plots into a single PDF

```{r}
myris <- iris

# "slow" way
my.answer <- data.frame(Species = NULL, Sepal.Length = NULL, Sepal.Width = NULL, Petal.Length = NULL, Petal.Width = NULL)
for(this.species in levels(myris$Species)) {
  myris.subset <- myris[myris$Species == this.species, ]
  sl <- mean(myris.subset$Sepal.Length)
  sw <- mean(myris.subset$Sepal.Width)
  pl <- mean(myris.subset$Petal.Length)
  pw <- mean(myris.subset$Petal.Width)
  this.answer <- data.frame(Species = this.species, Sepal.Length = sl, Sepal.Width = sw, Petal.Length = pl, Petal.Width = pw)
  my.answer <- rbind(my.answer, this.answer)
}

# "fast" way
my.answer <- data.frame(Species = rep(NA, length(levels(myris$Species))), Sepal.Length = NA, Sepal.Width = NA, Petal.Length = NA, Petal.Width = NA)
for(this.loop in 1:length(levels(myris$Species))) {
  this.species <- levels(myris$Species)[this.loop]
  myris.subset <- myris[myris$Species == this.species, ]
  sl <- mean(myris.subset$Sepal.Length)
  sw <- mean(myris.subset$Sepal.Width)
  pl <- mean(myris.subset$Petal.Length)
  pw <- mean(myris.subset$Petal.Width)
  my.answer[this.loop, ] <- list(this.species, sl, sw, pl, pw)
}


```


3. Make another copy of the loop, adding an `if` statement that will have the effect of skipping plants where `Petal.Width` is less than 1.

```{r}
# option 1 (no if)
my.answer <- data.frame(Species = rep(NA, length(levels(myris$Species))), Sepal.Length = NA, Sepal.Width = NA, Petal.Length = NA, Petal.Width = NA)
for(this.loop in 1:length(levels(myris$Species))) {
  this.species <- levels(myris$Species)[this.loop]
  myris.subset <- myris[myris$Species == this.species & myris$Petal.Width >= 1, ]
  sl <- mean(myris.subset$Sepal.Length)
  sw <- mean(myris.subset$Sepal.Width)
  pl <- mean(myris.subset$Petal.Length)
  pw <- mean(myris.subset$Petal.Width)
  my.answer[this.loop, ] <- list(this.species, sl, sw, pl, pw)
}

# option 2 (skipping whole species)
my.answer <- data.frame(Species = rep(NA, length(levels(myris$Species))), Sepal.Length = NA, Sepal.Width = NA, Petal.Length = NA, Petal.Width = NA)
for(this.loop in 1:length(levels(myris$Species))) {
  this.species <- levels(myris$Species)[this.loop]
  myris.subset <- myris[myris$Species == this.species & myris$Petal.Width >= 1, ]
  sl <- mean(myris.subset$Sepal.Length)
  sw <- mean(myris.subset$Sepal.Width)
  pl <- mean(myris.subset$Petal.Length)
  pw <- mean(myris.subset$Petal.Width)
  if(pw >= 1) {
    my.answer[this.loop, ] <- list(this.species, sl, sw, pl, pw)
  }
}


```

