---
title: "Day 6 Homework"
output: html_document
---

1. Think carefully about your data.  Think about something you would like to do multiple times, to subsets of your data.  

2. Write out (in plain English, in the notebook) what you are trying to do

3. Write some "pseudo-code" comments in an R chunk, as an outline

4. Write some (real) code that will do what you want ONCE, on one subset of the data.

5. Copy that code chunk, and convert it into a loop, which will do what you want over all the groups, concatenating the results.

```{r}
mynames <- c("a", "b", "c")

for(name in mynames) {
  print(name)
  
}


```

```{r}
library(reshape2)
setwd("~/R/Homework")
mySP <- read.csv("../r_programming_bsos_working/homework/dippolito/SPTimePeriods.csv")
head(mySP)
results <- table(mySP$T3)
head(results)

mySP$RelID <- 1:nrow(mySP)
mySP$RelID <- paste("Rel", 1:nrow(mySP), sep = "")
head(mySP)
mySP.melt <- melt(mySP, id.vars = c("RelID", "ID", "Relationship"), value.name = "relationship.type", variable.name = "time.period")
summary(mySP.melt)
head(mySP.melt)

results <- table(mySP.melt$relationship.type)
head(results)

my.answer <- data.frame(time.period = rep(NA, length(levels(mySP.melt$time.period))), relationship.type = NA)
for(this.loop in 1:length(levels(mySP.melt$time.period))) {
  this.tp <- levels(mySP.melt$time.period)[this.loop]
  mySP.subset <- mySP.melt[mySP.melt$time.period == this.tp, ]
  rt <- table(mySP.subset$relationship.type)
  this.answer <- data.frame(rt)
  this.answer <- dcast(this.answer, . ~ Var1, value.var = "Freq")[, -1]
  this.answer$time.period <- this.tp
  my.answer[this.loop, ] <- this.answer[1, ]
}
my.answer

```


6. Copy that loop again, and add a condition (`if`) statement to add some further control about ignoring some groups or some other conditional operation.
