---
title: "Day 4 Practice"
output: html_document
---

1. Read in the data base of phoneme pairs from Wedel, Kaplan, and Jackson (2013).  The relative path from the working directory of this project is: "scottdata/CognitionPaperModifiedData.csv"

```{r eval = TRUE}
# first one for running in RStudio, second path for knitting
#mydata <- read.csv("scottdata/CognitionPaperModifiedData.csv")
mydata <- read.csv("../scottdata/CognitionPaperModifiedData.csv")
summary(mydata)
```

2. Read in the data using the WRONG function.  What does it look like?

```{r eval = TRUE}
# first one for running in RStudio, second path for knitting
#mydata <- read.table("scottdata/CognitionPaperModifiedData.csv")
mydata <- read.table("../scottdata/CognitionPaperModifiedData.csv")
summary(mydata)

# Read in correctly again, for following code
#mydata <- read.csv("scottdata/CognitionPaperModifiedData.csv")
mydata <- read.csv("../scottdata/CognitionPaperModifiedData.csv")

```

3. Change some (or all) of the column names.  Provide a table that links original column names with the current/new column names.

```{r eval = TRUE}
mydata2 <- mydata
colnames(mydata2)
colnames(mydata2)[c(1, 3)] <- c("Language", "PhonemePair")
```

4. Locate NAs.  Remove all rows containing NAs.

```{r eval = TRUE}
# mydata2 <- mydata[!mydata$AllMinPairs == NA, ] # DOESN'T WORK

mydata2 <- mydata[!is.na(mydata$AllMinPairs), ]
summary(mydata2)

# more comprehensive
mydata3 <- mydata[complete.cases(mydata), ] # looks for any NAs in any column
summary(mydata3)

```


5. Recode the "Corpus" variable in the following way:
Am = American
Du = Dutch
Fr = French
Ge = German
HK = Cantonese
Ko = Korean
RP = British
Sl = Slovakian
Sp = Spanish

```{r}
old.levels <- levels(mydata$Corpus)
new.levels <- levels(mydata$Corpus) <- c("American", "Dutch", "French", "German", "Cantonese", "Korean", "British", "Slovakian", "Spanish")
print(data.frame(old.levels, new.levels))

```



6. Run a logistic regression, using `glm()`, where "Merged" status is the (binary) dependent variable, and "WithinMinPairs" is the predictor.
  - Hint #1: the basic usage of `glm()` is:
```{r eval=FALSE}
results <- glm(DV ~ predictor1 + predictor2 + ...  , 
               data = dataframe.object,
               family = "binomial")
```
  - Hint #2: you need the DV to be 0s and 1s, not a factor.
  
```{r eval= TRUE}
mydata$DV <- ifelse(mydata$Merged == "Merged", 1, 0)
mydata$DV2 <- as.numeric(mydata$Merged) - 1  # wrong order, but a possibility
results <- glm(DV ~ WithinMinPairs, data = mydata, family = "binomial")
```

  
7. Use `summary()` to display the results.

```{r}
summary(results)
```


8. Using INLINE code, report the estimate, standard error, and p-value for the WithinMinPairs coefficient.  This will look something like:

```{r eval = TRUE}
mysummary <- summary(results)
names(mysummary)
mysummary$coefficients
class(mysummary$coefficients)
mysummary$coefficients[2, 1]
est <- mysummary$coefficients["WithinMinPairs", "Estimate"]
std.err <- mysummary$coefficients["WithinMinPairs", "Std. Error"]
p.value <- mysummary$coefficients["WithinMinPairs", "Pr(>|z|)"]

```

The estimate for the WithinMinPairs coefficient was `r round(est, 4)`, with a standard error of `r round(std.err, 4)` and a p of `r round(p.value, 3)`.

EXTRA CREDIT: figure out how to display the above results with only a few decimal places

HINT: use `str()` to figure out how to extract these values.

