---
title: "Day 4 Homework"
output: html_document
---

1. Change the column names of your data to something easier to work with.  If you like your column names, change them into something else reasonable, just for practice.

2. List ALL the issues you can see where it looks like your data is being read in a weird way.  If your data looks perfect, find some way to mess it up :-)

```{r}
# colA and colB are "complementary" (one has NA where the other has value)
mydata <- data.frame(colA = c(NA, 1, 2, NA, NA, 78), colB = c(45, NA, NA, 2, 99, 23))
mydata
mydata$colC <- mydata$colA
mydata$colC[is.na(mydata$colC)]

mydata$colC[is.na(mydata$colA)] <- mydata$colB[is.na(mydata$colA)]
mydata

```

3. Pick one or two of the most egregious or problematic of the issues in #2 and fix them.

4. Check for NAs, and report where you see them (try to pinpoint where they are, not just "5 NAs in column seven".

```{r}
mydata <- data.frame(colA = c(NA, 1, 2, 3, NA, 78), colB = c(45, NA, NA, 2, 99, 23))
mydata
complete.cases(mydata)
mydata[!complete.cases(mydata), ]
which(!complete.cases(mydata))


```


5. Decide what you think you should do about the NAs, and say why.  If you can, try to implement this decision.



6. Remove any problematic rows and/or columns.  Say why you want to remove them.  If there aren't any problems, describe why you think there are no problems.

7. Re-code at least one factor.  Provide a table that shows the correspondence between old levels and new levels.



8. Run TWO DIFFERENT simple analyses or statistical tests, such as linear regression (`lm()`), logistic regression (`glm()`), correlation test (`cor.test()`), t-test (`t.test()`), or non-parametric tests (e.g., `wilcox.test()`).  For each of these:
  - Describe why you are doing this analysis, i.e., what question is it answering?
  - I won't judge you on statistical expertise!  (though I will make comments if I think I can be helpful)
  - Report some key statistics from the analysis, using inline code
  
  
```{r}
minpair <- read.csv("scottdata/CognitionPaperFinalData.csv")
head(minpair)
my.anova <- aov(WithinMinPairs ~ Merged, data = minpair)
my.summary <- summary(my.anova)
names(my.summary)
str(my.summary)
class(my.summary)
str(my.summary[[1]])
my.summary[[1]]$"F value"[1]
```

  