---
title: "Day 4 Lecture"
output: html_document
---

# Review?

# Data manipulation
  1. Simple column transformations
  2. Dealing with factors
  3. Reshaping data
  4. Aggregation

## Simple column transformations

```{r}
mysleep <- sleep
mysleep$extra2 <- mysleep$extra * 2
mysleep$extra3 <- log(mysleep$extra + 10)
mysleep
mysleep$extra2 <- mysleep$extra2 - 4
mysleep
mysleep$extra2 <- NULL  # deletes the column
mysleep
mysleep$extra3 <- NA
mysleep
mysleep$three <- 3
mysleep
mysleep$four <- c(1, 2, 3) # doesn't work
mysleep
```

## Dealing with factors
  - levels
  - droplevels
  - recoding
  - reordering
  - as.numeric vs. as.character
  - stringsAsFactors = argument to `read.table()`

### droplevels
```{r}
mysleep <- sleep
summary(mysleep)
levels(mysleep$ID)
summary(mysleep$ID)

mysleep.1 <- subset(mysleep, group == "1")
mysleep.1 <- subset(mysleep.1, !ID == "3")
# mysleep.1 <- subset(mysleep, group == 1 & !ID == 3)
# mysleep.1 <- mysleep[mysleep$group == 1 & !mysleep$ID == 3, ]
summary(mysleep.1$ID)
summary(mysleep.1$group)

mysleep.2 <- droplevels(mysleep.1)
summary(mysleep.2$ID)
summary(mysleep.2$group)
```

### recoding & reordering
  - relevel, reorder, factor, levels
```{r}
mysleep <- sleep
summary(mysleep)

levels(mysleep$group)
levels(mysleep$group) <- c("pre", "post")  # changes LABELS of the levels
levels(mysleep$group)
mysleep

# levels(mysleep$group) <- c("post", "pre")  # DOESN'T CHANGE ORDER!
# levels(mysleep$group)
# mysleep

mysleep$group <- relevel(mysleep$group, "post")  # changes order
levels(mysleep$group)
summary(mysleep)
mysleep

# avoid reorder() except for special cases

mysleep$ID <- factor(mysleep$ID, levels = c("2", "3", "1", "5", "10", "7", "9", "8", "4", "6"))
levels(mysleep$ID)
mysleep

mysleep$IDlevelnum <- as.numeric(mysleep$ID)
mysleep

```

 - `reorder()`
  - takes two vectors, a factor, and an "ordering" vector
  - the "ordering" vector needs to be:
    - the same length
    - numeric values corrsponding to the order of levels in the factor

## Reshaping data
Use the `reshape2` package!!!

```{r}
# install.packages("reshape2")
library(reshape2)
```

### "Long" vs. "wide" data
```{r}
mysleep <- sleep
mysleep
levels(mysleep$group) <- c("pre", "post")
mysleep.cast <- dcast(mysleep, ID ~ group, value.var = "extra")
mysleep.cast

myff <- french_fries
head(myff)
?melt.data.frame
myff.melt <- melt(myff, id.vars = 1:4)
myff.melt <- melt(myff, id.vars = c("time", "treatment", "subject", "rep"))
myff.melt <- melt(myff, measure.vars = 5:9)
myff.melt <- melt(myff, measure.vars = c("buttery", "potato", "grassy", "rancid", "painty"))
head(myff.melt)
summary(myff.melt)

myff.melt <- melt(myff, id.vars = 1:4, value.name = "rating", variable.name = "taste")
head(myff.melt)
summary(myff.melt)

recastff <- dcast(myff.melt, time + treatment + subject + rep ~ taste, 
                  value.var = "rating")
summary(recastff)
```

- Formula

## Aggregation (with the `reshape2` package)

```{r}
aggregateff <- dcast(myff.melt, treatment ~ taste, 
                  value.var = "rating", fun.aggregate = mean, na.rm = TRUE)

aggregateff
```

# Data cleaning

## A template for data cleaning
  1. Get the data to read in to R
  2. Fix column names
  3. Check all columns to make sure data is getting read in correctly
  4. Check NAs, decide what to do with them
  5. Remove excessive rows or other "defective" rows
  6. Re-code factors as needed
  
### Reading data
  - find a good `read.X()` function, where X = your type of data
    - `read.table()` is the basic one
    - `read.csv()` and `read.delim()` are convenience "wrappers" for `read.table()`, for CSV and tab-delimited files, respectively
    - `read.spss()` and `read.dta()` from the `foreign` package for SPSS and Stata files
  - for messier or more primitive kinds of files in weird formats, use `readLines()`, or `scan()` to pull in the data in a very raw format, which you will then need to format in R

### Fix column names

```{r}
mysleep <- sleep
oldnames <- colnames(mysleep)
mycolnames <- c("extraSleep", "experimentalGroup", "subjectID")
colnames(mysleep) <- mycolnames
data.frame(oldnames, mycolnames)

```

### Check all columns

### Check NAs
  - use summary()

### Remove rows/columns as needed

### Re-code factors as needed

# Complex objects

```{r}
myff <- french_fries
results <- lm(potato ~ treatment, data = myff)
results
summary(results)
plot(results)
class(results)
names(results)
plot(results$residuals)
results.summary <- summary(results)
names(results.summary)
results.summary$coefficients
str(results)
class(results.summary$coefficients)

str(summary(results))
?lm
```


## Looking at the structure of objects

## Extracting bits and pieces

