---
title: "Day 3 Homework"
output: html_document
---

The goal of this homework is to get practice with manipulating data.

1. Make both a "wide" and a "long" version of your data

- casting with a LOT of variables

```{r eval=FALSE}

mydata.wide <- dcast(mydata.long, [80 columns here] ~ split)

columns.to.keep <- colnames(mydata.long)[1:80]
columns.to.keep.pasted <- paste(columns.to.keep, collapse = " + ")
my.cast.formula <- as.formula(paste(columns.to.keep.pasted, " ~ split"))
class(my.cast.formula)  # "formula"
mydata.wide <- dcast(mydata.long, formula = my.cast.formula)


```

2. Transform a variable


3. Recode a factor

4. Add a new factor, where the levels are conditional on the values of some other variable. E.g., a factor of "high", "medium", and "low" based on thresholds for another variable.
```{r}



```


5. Reorder the levels of this new factor


6. Choose a factor that could be re-coded as multiple dimensions (recall the sepal/petal vs. length/width breakdown). If you don't have one, try to create one.  Now create two (or more) new factors that represent these dimensions.

7. Using `dcast`, aggregate some variable into "cell means" of some kind.

8.  Save all of the above in a well-documented R Markdown document.  Email me the .Rmd file (and if you use different or updated data from Day 2 Homework, send that to me, too).
