---
title: "Day 3 Practice"
output: html_document
---

**NOTE:**  this was postponed to Day 4

1. Make a copy of the `iris` data set in your workspace (i.e., assign the value of `iris` to a new name).

```{r}
myris <- iris
head(iris)
```

2. Reshape the data frame into "long" format.  
  - Can you still tell which measurements came from which plants?  If not, can you do something about that?
  - Are there better names for the `variable` and `value` columns?  Change them!

```{r}
myris$PlantID <- 1:nrow(myris)
myris$PlantID <- paste("plant", 1:nrow(myris), "ID", sep = "_")   # better!
head(myris)
myris.melt <- melt(myris, id.vars = c("PlantID", "Species"), 
                   value.name = "cm", variable.name = "measurement.type")
```

3. Create new columns that represent the part of the plant being measured (sepal, petal) and the dimension being measured (length, width).
  - Check to make sure it worked!
  
```{r}
# one way of doing it
myris.melt$Dimension <- "Width"
myris.melt$Dimension[myris.melt$measurement.type %in% c("Sepal.Length", "Petal.Length")] <- "Length"
xtabs(~ Dimension + measurement.type, data = myris.melt)

myris.melt$PlantPart <- "Petal"
myris.melt$PlantPart[myris.melt$measurement.type %in% c("Sepal.Length", "Sepal.Width")] <- "Sepal"
xtabs(~ Dimension + PlantPart + measurement.type, myris.melt)

# a better way: colsplit function from reshape2
myris.melt2 <- cbind(myris.melt, 
                     colsplit(myris.melt$measurement.type, pattern = "\\.", 
                              names = c("PlantPart", "Dimension")))
head(myris.melt2)
```

4. Create a data frame that leaves out the Sepal measurements

```{r}
myris.petals <- droplevels(myris.melt[myris.melt$PlantPart == "Petal", ])
summary(myris.petals)
```

5. Using the full data set, transform the measurement variable in some way
  - Create a new column to do it

```{r}
myris.melt$logcm <- log(myris.melt$cm)
summary(myris.melt)
```

6. Reshape the data so that there are two measurement columns, one for all Lengths and one for all Widths.

```{r}
myris.cast1 <- dcast(myris.melt, PlantID + Species + PlantPart ~ Dimension, value.var = "cm")
head(myris.cast1)
```

7. Using `dcast`, make a new data frame that aggregates over the flower parts (use mean or sum), but keeps the individual plants separate.

```{r}
myris.agg1 <- dcast(myris.melt, PlantID + Species ~ Dimension, value.var = "cm", fun.aggregate = mean)
head(myris.agg1)
myris.agg1 <- dcast(myris.melt, PlantID + Species + Dimension ~ ., value.var = "cm", fun.aggregate = mean)
head(myris.agg1)
```

8. Using `dcast`, make a new data frame that aggregates over flower parts and individual plants, getting "cell means" for species and dimension.

```{r}
myris.agg2 <- dcast(myris.melt, Species ~ Dimension, value.var = "cm", fun.aggregate = mean)
myris.agg2
myris.agg2 <- dcast(myris.melt, Species + Dimension ~ ., value.var = "cm", fun.aggregate = mean)
myris.agg2
```
