---
title: "Day 9 Lecture"
output: html_document
---

# HW 8 Review

# Special Topics
  - topic selection
  - make-up timing

# Functions
  - making a function
  - anonymous functions
  - environments
  - making functions available
  - debugging

## Function basics
  - basic syntax very simple!

```{r eval = FALSE}
my.function <- function(arg1, arg2, ...) {
  #FUNCTION BODY  
}

```

  - "anonymous" functions = "one-use" functions
  
```{r}
library(reshape2)
library(dplyr)
mysleep <- sleep
levels(mysleep$group) <- c("pre", "post")
mysleep.wide <- dcast(mysleep, ID ~ group, value.var = "extra")
mysleep.wide

double.max <- function(x) {
  output <- max(x) * 2
  output
}

mysleep.wide$newcol <- apply(mysleep.wide[, 2:3], 1, double.max)
mysleep.wide

# anonymous
mysleep.wide$newcol2 <- apply(mysleep.wide[, 2:3], 1, function(x) { max(x) * 2 })
mysleep.wide

# dplyr
mysleep.wide <- 
  mysleep.wide %>%
  rowwise() %>%
  mutate(newcol3 = double.max(c(pre, post))) # or newcol3 = max(pre, post) * 2
mysleep.wide

# dplyr with melt (good for lots of columns)
mysleep.wide %>%
  melt(measure.vars = 2:3) %>%
  group_by(ID) %>%
  mutate(newcol4 = double.max(value)) %>%
  dcast(ID + newcol + newcol2 + newcol3 + newcol4 ~ variable, value.var = "value")

# using `get()`
mysleep.wide %>%
  rowwise() %>%
  mutate(newcol5 = double.max(get(c("pre", "post")))) 


```

## Environments
  - R can look "up" through environments, but not "down"
  - the body of a function is an environment, nested under the environment where it is called

## Making functions available
  - in the notebook/script
  - `source()`
  - package!
    - some package-making tutorials:
      - http://hilaryparker.com/2014/04/29/writing-an-r-package-from-scratch/
      - http://r-pkgs.had.co.nz/
      - https://support.rstudio.com/hc/en-us/articles/200486488-Developing-Packages-with-RStudio
      - http://cran.r-project.org/doc/contrib/Leisch-CreatingPackages.pdf
      - and lots more

## Debugging
  - run as a non-function
  - insert "test" code
  - use `browser()`, `debug()`, and `undebug()`
  
```{r}
mysleep <- sleep

my.analysis <- function(x) {

}
```


# Practice

