---
title: "Day 9 Practice"
output: html_document
---

1. Use the following data:

```{r}
library(dplyr)
library(reshape2)
miris <- iris
set.seed(42)
for(col in 1:4) {
  miris[sample(nrow(miris), 20, replace = FALSE), col] <- NA
}

```

2. Write a function that will:
  - take a vector of numbers as input
  - return the log of the largest non-NA value
  - Try this out on a vector that you make up

```{r}
myfunc <- function(x) {
  x <- x[which(!is.na(x))]
  output <- log(max(x))
  output
}

myfunc(1:10)
myfunc(c(2, 10, 3, 4.1))
myfunc(c(2, 10, 3, 14.1))
myfunc(c(2, 10, NA, 14.1))


```


3. Run this function over rows of the data from #1

```{r}
miris$result <- apply(miris[, 1:4], 1, myfunc)

# dplyr way
miris <- miris %>% 
  rowwise() %>% 
  mutate(result2 = myfunc(c(Sepal.Length, Sepal.Width, Petal.Length, Petal.Width)))

# loop way
for(counter in 1:nrow(miris)) {
  result <- myfunc(miris[counter, 1:4])
  miris$newcolumn[counter] <- result
}


```


4. Write a function that will get the standard error of a vector of numbers. Use that function to get a standard error for each measure for each Species in the data from #1 (ignoring NAs)

```{r}
std.err <- function(x, ignore.na = TRUE) { 
  if(!ignore.na) {
    output <- sd(x)/sqrt(length(x)) 
  } else {
    numer <- sd(x, na.rm = TRUE)
    denom <- sqrt(length(which(!is.na(x))))
    output <- numer/denom
  }
  output
}

miris %>% 
  melt(id.vars = "Species") %>%
  group_by(Species, variable) %>%
  summarize(SE = std.err(value)) %>%
  dcast(Species ~ variable, value.var = "SE")

# loop
results.table <- data.frame(Species = levels(miris$Species), Sepal.Length = NA, Sepal.Width = NA, Petal.Length = NA, Petal.Width = NA)
for(species in levels(miris$Species)) {
  miris.sub <- filter(miris, Species == species)
  for(col in c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")) {
    this.answer <- std.err(miris.sub[, col])
    results.table[results.table$Species == species, col] <- this.answer
  }
}

```

