---
title: "Day 11 Lecture"
output: html_document
---

# HW 10 Review

# Final projects and grades

# Course recap, review, and suggestions for next steps

# Strings

  - regular expressions
  - grep, grepl, gsub
  - strsplit
  - colsplit

## Regular expressions

  - string *matching*
  - can match literal strings or more abstract *patterns*
  - see ?regex for help
  - regular expressions are NOT unique to R, so look for tips elsewhere, too
  - in R strings, `\` is the "escape" character, and for special regex expressions you usually need to double-escape (see examples)
  
Some handy characters to know:
  - `+` = match 1 or more of the previous character
  - `*` = match 0 or more of the previous character
  - `?` = the preceding item is optional (i.e., match 0 or 1 of the previous character). `?` appended to a repetition operator makes it "not greedy"
  - `[ ]` = match 1 of the set of things inside the bracket
  - `\w` = match a "word" character (i.e., letters and numbers).  NOTE: you need to double-escape this and the rest of the slashes below usually, like `\\w`
  - `\d` = match digits
  - `\s` = match a space character
  - capital-letter versions of the above (`\W`, `\D`, `\S`) are the *negations* of those
  - `\t` = match a "tab" character
  - `\n` = match a "newline" character
  - `^` = the "beginning edge" of a string
  - `$` = the "ending edge" of a string
  - {n} = the preceding character is matched n times

### basic matching

```{r}
response.types <- c("(1) strong dispreference", "(2) mild dispreference", "(3) no preference", "(4) mild preference", "(5) strong preference", "(9) don't know")
set.seed(89)
stringdata <- data.frame(ID = rep(paste("subject", 1:100), each = 4), 
                         condition = rep(c("small_slow", "small_fast", 
                                           "big_slow", "big_fast"), 100),
                         condition2 = rep(c("small.slow", "small.fast", 
                                           "big.slow", "big.fast"), 100),
                         condition3 = rep(c("smallslow", "smallfast", 
                                           "bigslow", "bigfast"), 100),
                         condition4 = rep(c("SmallSlow", "SmallFast", 
                                           "BigSlow", "BigFast"), 100),
                         response = sample(response.types, 400, replace = TRUE))

head(stringdata)
summary(stringdata)

x <- c("here", "is", "a", "string", "vector")
grep("v", x)
grep("i", x)

grepl("i", x)

# find rows with subject 12
grep("subject 12", stringdata$ID)
grep("12", stringdata$ID)
which(stringdata$ID == "subject 12")

# find all rows where the response was "mild"
grep("mild", stringdata$response)

# find all rows where condition has "small"
grep("small", stringdata$condition)

colnames(stringdata)

apply(stringdata, 2, function(x) grep("small", x))

x <- "hi there"
grepl("hi", x)
grepl("^hi", x)
grepl("there", x)
grepl("^there", x)
grepl("hi$", x)
grepl("there$", x)


```

### substitutions

```{r}

x <- "hello there"
gsub("ll", "", x)
gsub("l", "p", x)
gsub("e", "something else", x)
sub("e", "something else", x)

y <- "condition 58"
gsub("condition ", "", y)
as.numeric(gsub("condition ", "", y))

z <- c("c 58", "n 47", "o 83", "d 502", "number 3", "my num 42")
gsub("\\w ", "", z)
gsub("\\w+ ", "", z)
sub("\\w+ ", "", z)

# backreferences
gsub("\\w+ \\d+", "all", z) 
gsub("\\w+ \\d+", "all", z) 
gsub("(\\w+ )+\\d+", "all", z) 

gsub("(\\w+ )+(\\d+)", "\\2", z) 
gsub("(\\w+\\s)+(\\d+)", "\\2", z) 

# return just the first part of the condition variable
gsub("(\\w+)_(\\w+)", "\\1", stringdata$condition)

# return the ID column without the space
gsub("(\\w+)\\s(\\d+)", "\\1\\2", stringdata$ID)
gsub("\\s", "", stringdata$ID)

stringdata2 <- stringdata
stringdata2$ID <- as.character(stringdata2$ID)
stringdata2$ID[5:8] <- "s ubject 2"

gsub("(\\w+)\\s(\\d+)", "\\1\\2", stringdata2$ID)
gsub("\\s", "", stringdata2$ID)


```

### strsplit and colsplit

```{r}
library(reshape2)

x <- c("some string of words", "some other sentence", "yet another one")
strsplit(x, " ")
unlist(strsplit(x, " "))

unlist(strsplit(x, "\\w\\s\\w"))

x <- "some text <get rid of this> more text"
gsub("<(\\w+\\s*)+>\\s*", "", x)

# use strsplit to separate "small_slow"
x <- "small_slow"
unlist(strsplit(x, "_"))

# next, use strsplit to separate out two new columns from stringdata$condition column
x <- "small_slow"
unlist(strsplit(x, "_"))[1]
unlist(strsplit(x, "_"))[2]

stringdata$size <- 
  sapply(as.character(stringdata$condition), function(myarg) { unlist(strsplit(myarg, "_"))[1] })

stringdata$speed <- 
  sapply(as.character(stringdata$condition), function(x) { unlist(strsplit(x, "_"))[2] })

xtabs(~ size + speed + condition, stringdata) 

# using colsplit for the same thing
newcolumns <- colsplit(as.character(stringdata$condition), "_", names = c("size2", "speed2"))

stringdata <- cbind(stringdata, newcolumns)

xtabs(~ size2 + speed2 + condition, stringdata) 

# using colsplit for condition2
newcolumns <- colsplit(as.character(stringdata$condition2), "\\.", names = c("size3", "speed3"))

stringdata <- cbind(stringdata, newcolumns)

xtabs(~ size3 + speed3 + condition, stringdata) 

# using colsplit for condition4
newcolumns <- colsplit(as.character(stringdata$condition4), "[[:upper:]]", names = c("nothing", "size4", "speed4"))

x <- "abcdef"
gsub("abcdef", "X", x)
sub("[abcdef]", "X", x)

# \\w is equivalent to [0-9A-Za-z]
# [:upper:] is equivalent to A-Z  (on a standard English keyboard locale)
# [[:upper:]] is equivalent to [A-Z]
# [_\\.\\s]  will match ONE of underscore, dot, space

stringdata <- cbind(stringdata, newcolumns)

xtabs(~ size4 + speed4 + condition, stringdata) 

# using colsplit for condition3
x <- "bigfast"
y <- "smallslow"
# gsub("()()" ...
# gsub("((big)|(small)) ...
# gsub("((big)|(small))((fast)|(slow))" ...

gsub("((big)|(small))((fast)|(slow))", "\\1", x)
gsub("((big)|(small))((fast)|(slow))", "\\4", x)
gsub("((big)|(small))((fast)|(slow))", "\\1", y)
gsub("((big)|(small))((fast)|(slow))", "\\4", y)

stringdata$size5 <- gsub("((big)|(small))((fast)|(slow))", "\\1", stringdata$condition3)

stringdata$speed5 <- gsub("((big)|(small))((fast)|(slow))", "\\4", stringdata$condition3)
  
xtabs(~ size5 + speed5 + condition, stringdata) 

x <- c("one", "\\wo", "three")
gsub("\\\\", "", x)

sometext <- readLines("day11/trialtext.txt")



```


# Practice


