---
title: "Day 5 Lecture"
output: html_document
---

# Homework 3 Review

# Practice 4 Review

# Graphics overview

http://accidental-art.tumblr.com/

## Two systems

  - "base" graphics
  - "grid" graphics
  - See *R Graphics* by Paul Murrell for a very thorough "deep-dive" into these implementations:  https://www.stat.auckland.ac.nz/~paul/RG2e/

### Base graphics
  - more "primitive"
  - can also be much faster than grid
  - lots of isolated plot functions
  - lots of weird little parameters to worry about
  - some handy quick-plot functions from base graphics:
    - `plot()` : a generic function that will often produce something interesting, can do scatterplots with a formula `y ~ x`
    - `hist()` : histogram of a vector
    - `boxplot()` : boxplot, using a formula `variable ~ grouping.factor`

### Grid graphics
  - better for more "high-level" plots
  - much more complex underlying system (IMHO)
  - more powerful and extendable
  - the basis of extremely popular packages such as `lattice` and `ggplot2`

## Graphics packages

  - many packages will have some kind of graphics functions in them, for visualizing specific kinds of things
  - these will often be in base graphics, "pre-canned", etc.
  - the two best more general-purpose graphics packages are `lattice` and `ggplot2`
  - `ggplot2` is another Hadley Wickham project, and has the most documentation, examples, and is currently most widely used

# Using `ggplot2`

## A "Grammar of Graphics"
  - based on work by Leland Wilkinson, a "theory" of graphic display of data
  - Hadley implemented this framework as his dissertation
  - `ggplot2` is the overhauled version of the original implementation, and it is still under active development
  - the idea is to implement a general system for mapping data variables to *aesthetics* (Wilkinson's term), i.e., visual dimensions (x, y, color, etc.)
  - allows for high-level control of the structure of a plot, to allow for flexible, powerful visualizations
  
## Basic structure and concepts

  - unlike base graphics, the *value* of `ggplot2` objects is a complex object, and `print()`-ing them renders the graph.
  - the value of the object is built up by adding *layers*
  - start with a call to `ggplot()` that specifies the data and some mappings of variables to aesthetics
  - add "geoms" to actually render something
  - add other things for fine control over graphical parameters, etc.

## Graphing the `iris` package

```{r}
library(ggplot2)
myris <- iris
summary(myris)

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram()

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram(binwidth = .5)

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram(binwidth = .01)

mybaseplot <- ggplot(myris, aes(x = Sepal.Length))

mybaseplot + geom_histogram(binwidth = .2)

names(mybaseplot)
str(mybaseplot)

mybaseplot + geom_density(adjust = .5, fill = "grey")

```

```{r}
colors()
```

```{r}
ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + geom_point()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + geom_point(color = "red")

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point(aes(color = Species), size = 10) + 
  theme_minimal()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point(aes(color = Species, size = Petal.Width)) + 
  theme_minimal()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point() + 
  geom_smooth()

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width)) + 
  geom_point() + 
  geom_smooth(method = "lm")

ggplot(myris, aes(x = Sepal.Length, y = Sepal.Width, color = Species)) + 
  geom_point(size = 4) + 
  theme_minimal() +
  geom_smooth(method = "lm")


```

### Faceting

```{r}

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram() + 
  facet_wrap(~ Species)

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram() + 
  facet_grid(Species ~ .)

ggplot(myris, aes(x = Sepal.Length)) + geom_histogram() + 
  facet_grid(Species ~ .)

```

### Saving plots

```{r}
pdf("myplot.pdf", width = 10, height = 4)
ggplot(myris, aes(x = Sepal.Length)) + geom_histogram() + 
  facet_grid(Species ~ .)
dev.off()

?Devices
```


# Practice 5

# Review Practice 5

# Admin round-up

## Weekend homework

## Project proposals due!

## My flight schedule and Plan B
