---
title: "Day 8 Lecture"
output: html_document
---

# Vectorization overview
  - sometimes faster (not always)
  - sometimes easier to think about (not always)
  - split-apply-combine
  
# The `apply` family

## `apply`

- good for "row-wise" operations over data frames (or column-wise)
- give it only the columns/rows you want to operate on
- the `MARGIN` argument says whether it's row-wise (=1) or column-wise (=2)
- the `FUN` argument says what to do with each row/column-wise vector

```{r}
minpair <- read.csv("scottdata/CognitionPaperFinalData.csv")
head(minpair)
colnames(minpair)

minpair$allmean <- apply(minpair[, 10:17], MARGIN = 1, mean)
head(minpair)

```

## other members of the `apply` family
  - generally avoid, if possible!

### `lapply`
  - "list" apply
  - does something to each member of a list/vector
  - returns a list

### `tapply`
  - "table" apply
  - does something to data in "cells" by a combination of factors
  - returns a table

### `sapply`
  - "simple" apply
  - returns a vector
  - usually what you want instead of `lapply`

# A better way: `plyr` and `dplyr`
  - best for data frames
  
## `plyr`
  - the `plyr` package is older, slower, not under very active development
  - instantiates a "split-apply-combine" philosophy
  - handy paper here: http://www.jstatsoft.org/v40/i01
  - series of related functions
    - XYply, where X = "thing you give it" and Y = "thing it gives you back"
    - `ldply()` = you give it a list, it gives you back a data frame
    - `daply()` = you give it a data frame, it gives you back an array
    - etc.

## `dplyr`
  - the most common usage for `plyr` is `ddply()` -- give it a data frame and get a data frame back
  - the `dplyr` package takes this and runs with it
  - DON'T library both packages in the same session (unless you really need to and know what you're doing)
  - `dplyr` is implemented to be much faster
  - `dplyr` is the future...
  
### `dplyr` usage
  
  - "verbs": `mutate`, `filter`, `select`, `summarise` (or `summarize`), `arrange`
  - use `group_by()` to set the groups over which something (like `summarize` will apply)
  
```{r}



```

