---
title: "Day 8 Practice"
output: html_document
---

1. Start with this data:

```{r}
library(dplyr)
library(reshape2)
minpair <- read.csv("scottdata/CognitionPaperFinalData.csv")
```

2. Get the means of `WithinMinPairs` for each level of `Corpus`

```{r}
minpair %>%
  group_by(Corpus) %>%
  summarize(mean = mean(WithinMinPairs))

# this is equivalent!
summarize(group_by(minpair, Corpus), mean = mean(WithinMinPairs))

# you can mix
summarize(minpair %>% group_by(Corpus), mean = mean(WithinMinPairs))

```


3. Get cell means of `WithinMinPairs` by `Corpus` and `Merged`

```{r}
minpair %>%
  group_by(Corpus, Merged) %>%
  summarize(mean = mean(WithinMinPairs))
```

4. Melt down the last four columns, which are all frequency measures.  Get the min, max, and mean log values for all of these measures, by Corpus and Merged.
  - HINT: melt down the frequency columns first

```{r}
minpair %>% 
  melt(measure.vars = 14:17) %>%
  group_by(variable, Corpus, Merged) %>%
  summarize(minlog = min(log(value)), maxlog = max(log(value)), mean = mean(log(value)))
```

5. Do the same as 4, but only where HiSegTok is greater than the median value for each Corpus.

```{r}
step1 <- minpair %>% 
  group_by(Corpus) %>%
  summarize(medHST = median(HiSegTok))
step1

minpair.merged <- merge(step1, minpair)

minpair %>% 
  merge(step1) %>%
  filter(HiSegTok > medHST) %>%
  melt(measure.vars = 14:17) %>%
  group_by(variable, Corpus, Merged) %>%
  #summarize(minlog = min(log(value)), maxlog = max(log(value)), mean = mean(log(value))) %>%
  xtabs(~ medHST + Corpus, data = .)

minpair %>%
  group_by(Corpus) %>%
  mutate(medHST = median(HiSegTok)) %>%
  filter(HiSegTok > medHST) %>%
  melt(measure.vars = 14:17) %>%
  group_by(variable, Corpus, Merged) %>%
#  summarize(minlog = min(log(value)), maxlog = max(log(value)), mean = mean(log(value)))
  xtabs(~ medHST + Corpus, data = .)

```
