---
title: "Day 8 Lecture"
output: html_document
---

# Vectorization overview
  - sometimes faster (not always)
  - sometimes easier to think about (not always)
  - split-apply-combine

```{r}
x <- 1:10

x * 3

```


# The `apply` family

## `apply`

- good for "row-wise" operations over data frames (or column-wise)
- give it only the columns/rows you want to operate on
- the `MARGIN` argument says whether it's row-wise (=1) or column-wise (=2)
- the `FUN` argument says what to do with each row/column-wise vector

```{r}
minpair <- read.csv("scottdata/CognitionPaperFinalData.csv")
head(minpair)
colnames(minpair)

minpair$allmean <- apply(minpair[, 10:17], MARGIN = 1, FUN = mean)
head(minpair)

```

## other members of the `apply` family
  - generally avoid, if possible!

### `lapply`
  - "list" apply
  - does something to each member of a list/vector
  - returns a list

### `tapply`
  - "table" apply
  - does something to data in "cells" by a combination of factors
  - returns a table

### `sapply`
  - "simple" apply
  - returns a vector
  - usually what you want instead of `lapply`

# A better way: `plyr` and `dplyr`
  - best for data frames
  
## `plyr`
  - the `plyr` package is older, slower, not under very active development
  - instantiates a "split-apply-combine" philosophy
  - handy paper here: http://www.jstatsoft.org/v40/i01
  - series of related functions
    - XYply, where X = "thing you give it" and Y = "thing it gives you back"
    - `ldply()` = you give it a list, it gives you back a data frame
    - `daply()` = you give it a data frame, it gives you back an array
    - etc.

```{r eval = FALSE}
# library(plyr)
my.function <- (chunk)

results <- ddply(mydata, c("factor1", "factor1"), my.function)

```


## `dplyr`
  - the most common usage for `plyr` is `ddply()` -- give it a data frame and get a data frame back
  - the `dplyr` package takes this and runs with it
  - DON'T library both packages in the same session (unless you really need to and know what you're doing)
  - `dplyr` is implemented to be much faster
  - `dplyr` is the future...
  
### `dplyr` usage
  
  - "verbs": `mutate`, `filter`, `select`, `summarise` (or `summarize`), `arrange`
  - use `group_by()` to set the groups over which something (like `summarize` will apply)
  - `mutate` = "transform"
  - `filter` = "get some of the rows"
  - `select` = "get some of the columns"
  - `summarise` = boil things down to some kind of summary
  
```{r}
library(dplyr)
minpair <- mutate(minpair, newcolumn = log(HiSegTok), anothercol = log(HiSegType))
head(minpair)
# the same as
minpair$newcolumn <- log(minpair$HiSegTok)

minpair.subset <- filter(minpair, Corpus != "Am", log(LoSegType) > -5) 
summary(minpair)
summary(minpair.subset)

minpair.column.subset <- select(minpair, Corpus, Merged)
head(minpair.column.subset)
# only keep a stretch of columns
minpair.column.subset <- select(minpair, AllMinPairs:WithinMinPairs)
head(minpair.column.subset)
# pull out a stretch of columnns
minpair.column.subset <- select(minpair, -(AllMinPairs:WithinMinPairs))
head(minpair.column.subset)
minpair.column.subset <- select(minpair, Corpus:Merged, AllMinPairs:WithinMinPairs)
head(minpair.column.subset)

summarize(minpair, mymean = mean(HiSegTok), anothermean = mean(LoSegTok))
# same as
data.frame(mymean = mean(minpair$HiSegTok), anothermean = mean(minpair$LoSegTok))


# group_by()
minpair.grouped <- group_by(minpair, Corpus)
nrow(minpair)
nrow(minpair.grouped)
head(minpair)
head(minpair.grouped)
# clean-up (sometimes, but rarely, needed)
minpair.grouped.clean <- as.data.frame(minpair.grouped)
head(minpair.grouped.clean)

summarize(minpair.grouped, means = mean(HiSegTok))

x <- 1:10
y <- log(x)
z <- sd(y)

sd(log(1:10))

1:10 %>% 
  log %>% 
  sd

minpair %>% 
  filter(WithinMinPairs > 2) %>%
  group_by(Corpus) %>%
  summarize(mean1 = mean(HiSegTok), mean2 = mean(LoSegTok))

minpair %>% 
  group_by(Corpus, Merged) %>%
  summarize(mean1 = mean(HiSegTok), mean2 = mean(LoSegTok))
  
minpair %>% 
  group_by(Corpus, Merged) %>%
  summarize(mean1 = mean(HiSegTok), mean2 = mean(LoSegTok)) %>%
  summarize(mean1 = mean(mean1), mean2 = mean(mean2))

# You can mix these different modes of writing code
sd(1:10 %>% log)



```

